---
title: "**Predicting Wine Quality from Climate Data**: A Bordeaux Case Study"
author: "Dargie Tsegay Berhe <br> dtb2434@my.utexas.edu <br> tsegadar@yahoo.com <br> The University of Texas at Austin, USA"
date: "`r Sys.Date()`" 
output: 
  html_document:
    theme: readable
    toc: true
---

# 1. Introduction
This project analyzes how historical climate variables—specifically temperature and rainfall—impact the quality and price of Bordeaux wines. Leveraging domain knowledge in viticulture, I process historical weather data and apply statistical modeling to predict vintage quality before the wine is even bottled.

# 2. Environment Setup
```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(rsample)
```

#3. Data Ingestion & Wrangling

```{r}
# Pulling the dataset directly from a public repository
data_url <- "https://raw.githubusercontent.com/egarpor/handy/master/datasets/wine.csv"
bordeaux_raw <- read_csv(data_url)

bordeaux_clean <- bordeaux_raw |>
  drop_na(Price) |>
  mutate(
    Age = max(Year) - Year,
    # Flag drought conditions during harvest (a known viticultural risk factor)
    DryHarvest = if_else(HarvestRain < 100, "Yes", "No")
  ) |> 
  select(Year, Price, AGST, WinterRain, HarvestRain, Age, DryHarvest)
```

# 4. Rigorous Validation: Train/Test Split

```{r}
set.seed(2026) 

# Reserve 80% of the data to train the model, hide 20% to test it later
wine_split <- initial_split(bordeaux_clean, prop = 0.80)
train_data <- training(wine_split)
test_data  <- testing(wine_split)
```

# 5. Model Training and Evaluation
```{r}
# Train a Multiple Linear Regression model
final_wine_model <- lm(Price ~ AGST + WinterRain + HarvestRain + Age + DryHarvest, data = train_data)

# Test the model on unseen data
test_predictions <- test_data |>
  mutate(Predicted_Price = predict(final_wine_model, newdata = test_data))

# Calculate Root Mean Squared Error (RMSE)
test_results <- test_predictions |>
  mutate(
    Error = Price - Predicted_Price,
    Squared_Error = Error^2             
  ) |>
  summarize(RMSE = sqrt(mean(Squared_Error)))

print(test_results)
```
